---
title: "p8105_hw2_hb2699"
author: "Han Bao"
date: "`r Sys.Date()`"
output: html_document
---
# `r "1. NYC data"`
```{r}
library(tidyverse)
```
```{r}
NYC_data = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))

NYC_data = janitor::clean_names(NYC_data)
```


```{r}
NYC_data2 = select(NYC_data,line, station_name, station_latitude, station_longitude,entry, vending, entrance_type,ada, route1:route11)
NYC_data$entry<-ifelse(NYC_data$entry=="Yes",1,0)
skimr::skim(NYC_data)
#git push -f origin main
```



#### `r "variables description"`
This dataset contains line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.The steps I did so far is standardizing the name of the variables and changing the entry variable from yes/no to 1/0. The dimension of the dataset is 1868*32. I would say the dataset is not tidy enough since the route varibale takes 11 columns.

#### `r "Questions"`
1.We can see from the codes below that there are 465 distinct stations in this dataset.

```{r}
library(dplyr)
NYC_data2$dis_station <- paste(NYC_data2$line, NYC_data2$station_name)
distinct(NYC_data2, dis_station)
```
2.From the code below we can see that there are 468 compliant ada.
```{r}
sum(NYC_data2$ada, na.rm = TRUE)
```
3.We can see from the below code that 9.797% of station entrances / exits are without vending allow entrance.
```{r}
filter(NYC_data2, vending =="NO")
filter(NYC_data2, vending =="YES")
183/(183+1685)
```
```{r}
is.character(NYC_data2$route1)
is.character(NYC_data2$route8)
```

```{r}
NYC_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYC_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
# `r "2. TrashWheel data"`
```{r}
mr_trash_df= readxl::read_excel("./Trash Wheel Collection Data.xlsx", 
                                 sheet = "Mr. Trash Wheel",
                                 range = cellranger::cell_cols("A:N"))
trash = janitor::clean_names(mr_trash_df)
trash$sports_balls = as.integer(trash$sports_balls)
```
```{r}
mr_trash_df2= readxl::read_excel("./Trash Wheel Collection Data.xlsx", 
                                 sheet = "Professor Trash Wheel",
                                 range = cellranger::cell_cols("A:N"))
trash2 = janitor::clean_names(mr_trash_df2)
trash$fromM <- c("Mr")
trash2$fromF <- c("Prof")
#left-join the two datasets
trash3 = 
  left_join(trash, trash2, by = "dumpster")
#the total weight of trash collected by Professor Trash Wheel
sum(trash2$weight_tons)
#the total number of sports balls collected by Mr. Trash Wheel in 2020
sum(trash[which(trash$year=='2020'),"sports_balls"])
```

The number of observations in the resulting dataset 548. Key vatiables include: the date of each collection(e.g. date.x), the source of the collection(e.g. fromM), the type of the collection(e.g. plastic_bottles.x), weight of the collection(weight_tons.y) and etc. The total weight of trash collected by Professor Trash Wheel is 380.24 tons.The total number of sports balls collected by Mr. Trash Wheel in 2020 is 856.

# `r "3. 538"`
```{r}
#read in the files
snp = read_csv(file = "./snp.csv")
pols = read_csv(file = "./pols-month.csv")
une = read_csv(file = "./unemployment.csv")
```
```{r}
#separate the date into three variables
pols = separate(pols, col = mon, into = c('year','month','day'), sep = '-')
```
```{r}
#replace month number with month name
pols$month = as.numeric(pols$month)
pols$month <- month.abb[pols$month]
 
```
```{r}
#replace varibales and drop the useless variables
pols =mutate(pols, president = ifelse(prez_gop == "0","dem","gop"))
pols = select(pols, -prez_dem, -prez_gop,-day)
pols = janitor::clean_names(pols)
```
```{r}
#do the same process to data in snp.csv
snp = janitor::clean_names(snp)
snp = mutate(snp, date = lubridate::parse_date_time2(date,orders ="mdy", cutoff_2000 = 23))
```
```{r}
#separate the date into three variables
snp = separate(snp, col = date, into = c('year','month','day'), sep = '-')
```
```{r}
#replace month number with month name
snp$month = as.numeric(snp$month)
snp$month <- month.abb[snp$month]
snp = select(snp,-day)
#arrange the dataset according to year and month
snp = arrange(snp,year,month)

 
```
```{r}
#merge 
une = pivot_longer(une,Jan:Dec, names_to = "month", values_to = "unemployment_rate") 
une = janitor:: clean_names(une)
```
```{r}
# merge 
merge1 = left_join(pols, snp, by = "year","month")
#colnames(merge1)[3] <- "month"
#merge2 = left_Join(merge1, une, by = c("year","month"))
```







