---
title: "p8105_hw2_hb2699"
author: "Han Bao"
date: "`r Sys.Date()`"
output: html_document
---
# `r "1. NYC data"`
```{r}
library(tidyverse)
```
```{r}
NYC_data = read_csv(file = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))

NYC_data = janitor::clean_names(NYC_data)
```


```{r}
NYC_data2 = select(NYC_data,line, station_name, station_latitude, station_longitude,entry, vending, entrance_type,ada, route1:route11)
NYC_data$entry<-ifelse(NYC_data$entry=="Yes",1,0)
skimr::skim(NYC_data)
#git push -f origin main
```



#### `r "variables description"`
This dataset contains line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.The steps I did so far is standardizing the name of the variables and changing the entry variable from yes/no to 1/0. The dimension of the dataset is 1868*32. I would say the dataset is not tidy enough since the route varibale takes 11 columns.

#### `r "Questions"`
1.We can see from the codes below that there are 465 distinct stations in this dataset.

```{r}
library(dplyr)
NYC_data2$dis_station <- paste(NYC_data2$line, NYC_data2$station_name)
distinct(NYC_data2, dis_station)
```
2.From the code below we can see that there are 468 compliant ada.
```{r}
sum(NYC_data2$ada, na.rm = TRUE)
```
3.We can see from the below code that 9.797% of station entrances / exits are without vending allow entrance.
```{r}
filter(NYC_data2, vending =="NO")
filter(NYC_data2, vending =="YES")
183/(183+1685)
```
```{r}
is.character(NYC_data2$route1)
is.character(NYC_data2$route8)
```

```{r}
NYC_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

NYC_data %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
# `r "2. TrashWheel data"`
```{r}
mr_trash_df= readxl::read_excel("./Trash Wheel Collection Data.xlsx", 
                                 sheet = "Mr. Trash Wheel",
                                 range = cellranger::cell_cols("A:N"))
trash = janitor::clean_names(mr_trash_df)
trash$sports_balls = as.integer(trash$sports_balls)
```



